---
title: "TBI Onboard Data Preliminary Results"
author: "Yaxuan (Sean) Zhang"
date: "2023-08-24"
output: html_document
---

<font size="4">
\
• Who Rides Transit in 2022\
• Transit Trends before and after Covid\
• New Insights From 2022 Data\
\
</font>

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(data.table)
library(dplyr)
library(plotly)
library(councilR)
library(stringr)
library(readxl)
# Read 2016 tbi ------
svy_2016 <- read.csv('data/TravelBehaviorInventory2016TransitOnBoardSurvey.csv') %>% as.data.table()
svy_2016 <- svy_2016[DATE!='Dummy Record']

# Read 2022 tbi ------
svy_2022 <- read.csv('data/OD_2022.csv') %>% as.data.table()

# Format tbi data -------------
source("R/format/d_trip_type.R")
source("R/format/d_student.R")
source("R/format/d_race.R")
source("R/format/d_income.R")
source("R/format/d_hour_from_time.R")
source("R/format/d_gender.R")
source("R/format/d_employment.R")
source("R/format/d_disability.R")
source("R/format/d_age.R")
source("R/format/d_poverty.R")
source("R/format/d_drive.R")
source("R/format/d_work_location.R")
source("R/format/d_access_mode.R")

# Save matched route data--------------
svy_2016[, Route := fcase(
  Route == "19",
  "CLINE",
  rep_len(TRUE, nrow(svy_2016)),
  Route
)]

svy_2022[, ROUTE := fcase(
  ROUTE == "612",
  "12",
  ROUTE == "901",
  "Blue",
  ROUTE == "902",
  "Green",
  ROUTE == "903",
  "Red",
  ROUTE == "904",
  "Orange",
  ROUTE == "921",
  "ALINE",
  ROUTE == "923",
  "CLINE",
  rep_len(TRUE, nrow(svy_2022)),
  ROUTE
)]

svy_2016 <- svy_2016[Route %in% unique(svy_2022$ROUTE)]  #120 routes matched
svy_2022p <- svy_2022[ROUTE %in% unique(svy_2016$Route)]  #120 routes matched

# Read and format acs data ------
source("R/format/d_ACS.R")
# Table funs ----
source("R/format/fun_gettable.R")
# Fonts ----
library(sysfonts)
library(showtext)

font_add("HelveticaNeueLT Std Lt", "Font/HelveticaNeueLTStd-Lt.otf")
font_add( "HelveticaNeueLT Std Cn", "Font/HelveticaNeueLTStd-MdCn.otf")
font_add("Arial Narrow", "Font/ARIALN.ttf")
font_add("Arial Narrow Italic",
         regular = "Font/ARIALN.ttf",
         italic = "Font/ARIALNI.ttf")
showtext::showtext_auto()

```

### Who Rides Transit in 2022?

 <font size="4"> We explore demographic characteristics of public transit users (136 routes in total) in 2022. We also compare the demographic distribution to the general population distribution from American Community Survey (ACS).\
 </font>
 
#### **by Gender**
```{r fig_gender, echo=FALSE}
source("R/figure/fig_gender.R")
fig
```

#### **by Disability**
```{r fig_disability, echo=FALSE}
source("R/figure/fig_disability.R")
fig
```
#### **by Household Income/ Poverty Level**
```{r fig_poverty, echo=FALSE}
source("R/figure/fig_poverty.R")
fig
```

#### **by Vehicle Ownership**
```{r fig_car, echo=FALSE}
source("R/figure/fig_car.R")
fig
```
#### **by Race**
```{r fig_race, echo=FALSE, fig.dim = c(8, 7.2)}
source("R/figure/fig_race.R")
fig
```
#### **by Age**
```{r fig_age, echo=FALSE}
source("R/figure/fig_age.R")
fig
```

#### **by Employment**
```{r fig_employment, echo=FALSE}
source("R/figure/fig_employment.R")
fig
```
### Transit Trends before and after Covid

<font size="4"> We compare trips with same routes (120 routes in total) in 2016 and 2022. 
The results show that of the 317,393 daily trips taken on these routes in 2016. 141,948 trips, or 45%, were retained in 2022.\
</font>

#### **Trips lost and retained: by Gender**
```{r fig_gender_retain, echo=FALSE}
source("R/figure/fig_gender_retain.R")
fig
```
#### **Trips lost and retained: by Gender & Employment**
```{r fig_gender_x_employment_retain, echo=FALSE}
source("R/figure/fig_gender_x_employment_retain.R")
fig
```

#### **Trips lost and retained: by Trip type**
```{r fig_trip_type_retain, echo=FALSE}
source("R/figure/fig_trip_type_retain.R")
fig
```
#### **Trips lost and retained: by Gender & Trip Type**
```{r fig_gender_x_trip_type_retain, echo=FALSE}
source("R/figure/fig_gender_x_trip_type_retain.R")
fig
```
#### **Trips lost and retained: by Disability & Age**
```{r fig_age_x_disability_retain, echo=FALSE}
source("R/figure/fig_age_x_disability_retain.R")
fig
```

#### **Trips lost and retained: by Household Income & Student**
```{r fig_income_x_student_retain, echo=FALSE}
source("R/figure/fig_income_x_student_retain.R")
fig
```
### New Insights From 2022 Data
#### **More comprehensive gender definition** 
```{r fig_gender_new_categories, echo=FALSE}
source("R/figure/fig_gender_new_categories.R")
fig
```

#### **Why do different gender groups travel?**
```{r fig_gender_trip_type_new, echo=FALSE}
source("R/figure/fig_gender_trip_type_new.R")
fig
```

#### **When do different gender groups travel during the day?**
```{r fig_gender_trip_time, echo=FALSE}
source("R/figure/fig_gender_trip_time.R")
fig
```

#### **Alternate Mode **
```{r fig_drive_trip_type, echo=FALSE}
source("R/figure/fig_drive_trip_type.R")
fig
```
#### **How people travel in work-from-home era**
```{r fig_work_trip_type, echo=FALSE}
source("R/figure/fig_work_trip_type.R")
fig
```

